HTW Model

### 2024-02-15 Model/htw_model.qmd
render after big updates to HTW ABC modeling

---
title: HTW Model
author: Thomas Gorman
date: "`r Sys.Date()`"
code-fold: true
code-tools: true
execute: 
  warning: false
  eval: true
---

@@

The general ABC procedure is to 
1) define a prior distribution over model parameters.
2) sample candidate parameter values, $\theta^*$, from the prior.
3) Use $\theta^*$ to generate a simulated dataset, $Data_{sim}$.
4) Compute a measure of discrepancy between the simulated and observed datasets, $discrepancy$($Data_{sim}$, $Data_{obs}$). 
5) Accept $\theta^*$ if the discrepancy is less than the tolerance threshold, $\epsilon$, otherwise reject $\theta^*$.
6) Repeat until desired number of posterior samples are obtained.

Although simple in the abstract, implementations of ABC require researchers to make a number of non-trivial decisions as to i) the discrepancy function between observed and simulated data, ii) whether to compute the discrepancy between trial level data, or a summary statistic of the datasets, iii) the value of the minimum tolerance $\epsilon$. For the present work, we follow the guidelines from published ABC tutorials [@turnerTutorialApproximateBayesian2012; @farrellComputationalModelingCognition2018]. For the test stage, we summarized datasets with mean velocity of each band in the observed dataset as $V_{obs}^{(k)}$ and in the simulated dataset as $V_{sim}^{(k)}$, where $k \in \{1, 2, 3, 4, 5, 6\}$ represents each of the six velocity bands. For computing the discrepancy between datasets in the training stage, we aggregated training trials into three equally sized blocks (separately for each velocity band in the case of the varied group). After obtaining the summary statistics of the simulated and observed datasets, the discrepancy was computed as the mean of the absolute difference between simulated and observed datasets (#eq-discrep-test and #eq-discrep-train).

$$
discrepancy_{Test}(Data_{sim}, Data_{obs}) = \frac{1}{6} \sum_{k=1}^{6} |V_{obs}^{(k)} - V_{sim}^{(k)}|
$$  {#eq-discrep-test}

$$
\begin{aligned} \\
discrepancy_{Train,constant}(Data_{sim}, Data_{obs}) = \frac{1}{N_{blocks}} \sum_{j=1}^{N_{blocks}} |V_{obs,constant}^{(j)} - V_{sim,constant}^{(j)}| \\ \\
discrepancy_{Train,varied}(Data_{sim}, Data_{obs}) = \frac{1}{N_{blocks} \times 3} \sum_{j=1}^{N_{blocks}} \sum_{k=1}^{3} |V_{obs,varied}^{(j,k)} - V_{sim,varied}^{(j,k)}|
\end{aligned}
$$  {#eq-discrep-train}

@@

```{r}
pacman::p_load(dplyr,purrr,tidyr,ggplot2, data.table, here, patchwork, conflicted, 
               stringr,future,furrr, knitr, reactable, flextable,ggstance, htmltools)
conflict_prefer_all("dplyr", quiet = TRUE)
options(scipen = 999)
walk(c("Display_Functions","fun_alm","fun_indv_fit","fun_model"), ~ source(here::here(paste0("Functions/", .x, ".R"))))
ds <- readRDS(here::here("data/e1_md_11-06-23.rds"))  |> as.data.table()
nbins <- 3

@@

input_layer <<- output_layer <<-  c(100,350,600,800,1000,1200)
ids2 <- c(1,66,36)
ids3 <- c(20,71,101,4,76,192)
idsBad <- c(76,192, 101)

#file_name <- "n_iter_100_ntry_400_3247"
file_name <- "n_iter_200_ntry_300_5354"

ind_fits <- map(list.files(here(paste0('data/abc_reject/'),file_name),full.names=TRUE), readRDS)
# ind_fits <- map(list.files(here('data/abc_reject/n_iter_2000_ntry_10_2918'),full.names=TRUE), readRDS)
ind_fits_df <- ind_fits |> map(~list(dat=.x[[1]], Model = .x[["Model"]], Fit_Method=.x[["Fit_Method"]]))
ind_fits_df <- ind_fits_df |> map(~rbindlist(.x$dat) |> mutate(Model = .x$Model, Fit_Method = .x$Fit_Method)) |> rbindlist()

# avg_samples <- ind_fits_df |> group_by(id,condit,Model,Fit_Method) |> 
#   summarise(inc=max(inc_count),iter_sum=sum(iter_count))
# mean(avg_samples$iter_sum)
# mean(avg_samples$iter_sum)
# min(avg_samples$iter_sum)
# max(avg_samples$iter_sum)
# sum(avg_samples$iter_sum)
# sum(avg_samples$iter_sum)
# avg_samples |> group_by(Model,Fit_Method) |> summarise(mean(inc),mean(iter_sum))

@@

post_dat_avg <- post_dat |> group_by(id, condit, Model, Fit_Method, x, c, lr, rank) |> 
  summarise(y = mean(y), pred = mean(pred), error = y - pred) |> as.data.table()

@@

# organize training data predictions
pd_train <- nestSbjModelFit |> mutate(pp=furrr::future_pmap(list(id,Model,Fit_Method,data), ~{
   generate_data(..2, ..4 |> mutate(id=..1), ds, num_samples = 1, return_dat="train_data")
   })) |>
  select(Fit_Method,pp,-data) |>
  unnest(pp) |> as.data.table() |> filter(expMode2=="Train")
plan(sequential)

@@

bestTestEXAM <- post_dat_l |> 
  filter(Fit_Method=="Test", rank==1, Resp=="EXAM") |> 
  group_by(id,condit,Model,Fit_Method) |>
  summarise(mae=mean(abs(error))) |> arrange(mae) |> ungroup() |>
  mutate(group_rank = row_number(),.by=c(condit))

bestTestALM<- post_dat_l |> 
  filter(Fit_Method=="Test", rank==1, Resp=="ALM") |> 
  group_by(id,condit,Model,Fit_Method) |>
  summarise(mae=mean(abs(error))) |> arrange(mae) |> ungroup() |>
  mutate(group_rank = row_number(),.by=c(condit))

bestTrainALM<- post_dat_l |> 
  filter(Fit_Method=="Train", rank==1, Resp=="ALM") |> 
  group_by(id,condit,Model,Fit_Method) |>
  summarise(mae=mean(abs(error))) |> arrange(mae) |> ungroup() |>
  mutate(group_rank = row_number(),.by=c(condit))

bestTrainEXAM<- post_dat_l |> 
  filter(Fit_Method=="Train", rank==1, Resp=="EXAM") |> 
  group_by(id,condit,Model,Fit_Method) |>
  summarise(mae=mean(abs(error))) |> arrange(mae) |> ungroup() |>
  mutate(group_rank = row_number(),.by=c(condit))

bestTtEXAM <- post_dat_l |> 
  filter(Fit_Method=="Test_Train", rank==1, Resp=="EXAM") |> 
  group_by(id,condit,Model,Fit_Method) |>
  summarise(mae=mean(abs(error))) |> arrange(mae) |> ungroup() |>
  mutate(group_rank = row_number(),.by=c(condit))

@@

```{r}
#| eval: true
#| label: tbl-htw-modelError
#| tbl-cap: "Mean model errors predicting testing data, aggregated over all participants and velocity bands. Note that Fit Method refers to how model parameters were optimized, while error values reflect mean absolute error for the 6 testing bands"

post_tabs <- abc_tables(post_dat,post_dat_l)
post_tabs$agg_pred_full |> flextable::tabulator(rows=c("Fit_Method","Model"), columns=c("condit"), 
                       `ME` = as_paragraph(mean_error)) |> as_flextable()

@@

```{r}
#| label: fig-htw-residuals
#| fig-cap: Model residuals for each combination of training condition, fit method, and model. Residuals reflect the difference between observed and predicted values. Lower values indicate better model fit. Note that y axes are scaled differently between facets.
#| fig-subcap: ["Model Residuals - training data", "Model Residuals - testing data"]
#| fig-height: 10
#| fig-width: 11

pd_train |> group_by(id,condit,Model,Fit_Method, Block) |> 
  summarise(y = mean(y), pred = mean(pred), error = y - pred) |>
  ggplot(aes(x = Block, y = abs(error), fill=Model)) + 
  stat_bar + 
  facet_wrap(condit~Fit_Method, scales="free") +
  labs(title="Model Residuals - Training", y="Residual Error", x= "Training Block")

post_dat_avg |> 
  #filter(Fit_Method=="Test_Train") |> 
  ggplot(aes(x=abs(error),y=as.factor(x),col=Model)) + 
  stat_pointinterval(position=position_dodge(.2)) + 
  ggh4x::facet_nested_wrap(condit~Fit_Method, axes = "all",ncol=3,scale="free") +
  labs(y="Velocity Band", x="Residual Error")

# post_dat_avg |> 
#   #filter(Fit_Method=="Test_Train") |> 
#   ggplot(aes(x=(error),y=as.factor(x),col=Model)) + 
#   stat_pointinterval(position=position_dodge(.2)) + 
#   ggh4x::facet_nested_wrap(condit~Fit_Method, axes = "all",ncol=3,scale="free") +
#   labs(y="Velocity Band", x="Residual Error")

```
```{r}
#| label: fig-htw-obs-pred
#| fig-cap: Full posterior predictive distributions vs. observed data from participants.
#| fig-height: 10
#| fig-width: 11
post_dat_l |> 
  #filter(Fit_Method=="Test_Train") |> 
  ggplot(aes(x=val,y=as.factor(x),col=Resp)) + 
  stat_pointinterval(position=position_dodge(.2)) + 
  ggh4x::facet_nested_wrap(condit~Fit_Method, axes = "all",ncol=3,scale="free") +
  labs(y="Target Velocity Band (lower bound)", x="Velocity")

```{r}
#| label: fig-htw-indv-pred2
#| fig-cap: Individual predictions
#| fig-height: 15
#| fig-width: 12

post_dat_l |> 
  filter(Fit_Method=="Test_Train",x==100) |> 
  group_by(x) |>
  mutate(id = reorder(id, val)) |>
  ggplot(aes(x=val,y=id,col=Resp)) + 
  stat_pointinterval(position=position_dodge(.7)) + 
  ggh4x::facet_grid2(condit~x, axes = "all",scales="free", independent = "y") +
  labs(y="Target Velocity Band (lower bound)", x="Velocity")

post_dat_l |> 
  filter(Fit_Method=="Test_Train", id %in% ids2) |> 
  group_by(x) |>
  mutate(id = reorder(id, val)) |>
  ggplot(aes(x=val,y=as.factor(x),col=Resp)) + 
  stat_pointinterval(position=position_dodge(.2)) + 
  ggh4x::facet_nested_wrap(condit~id, axes = "all",scales="free_x") +
  labs(y="Target Velocity Band (lower bound)", x="Velocity")

post_dat_l |> 
  filter(Fit_Method=="Test_Train",x==100) |> 
  group_by(x) |>
  mutate(id = reorder(id, val)) |>
  ggplot(aes(x=val,y=id,col=Resp)) + 
  stat_pointinterval(position=position_dodge(.7)) + 
  ggh4x::facet_nested_wrap(condit~x, axes = "all",scales="free") +
  labs(y="Target Velocity Band (lower bound)", x="Velocity")

# post_dat_avg |> 
#   mutate(id = reorder(id, abs(error))) |>
#   #filter(Fit_Method=="Test_Train") |> 
#   ggplot(aes(x=abs(error),y=id,col=Model)) + 
#   stat_pointinterval(position=position_dodge(.2)) + 
#   ggh4x::facet_nested_wrap(condit~Fit_Method, axes = "all",ncol=3,scale="free") +
#   labs(y="Velocity Band", x="Residual Error")
# 
# 
# post_dat_avg |> 
#   filter(Fit_Method=="Test_Train") |> 
#   group_by(x) |>
#   mutate(id = reorder(id, abs(error))) |>
#   ggplot(aes(x=abs(error),y=id,col=Model)) + 
#   stat_pointinterval(position=position_dodge(.2)) + 
#   ggh4x::facet_nested_wrap(condit~x, axes = "all",ncol=6,scale="free") +
#   labs(y="Velocity Band", x="Residual Error")
# 
# 
# post_dat_avg |> 
#   filter(Fit_Method=="Test_Train") |> 
#   group_by(x) |>
#   mutate(id = reorder(id, (error))) |>
#   ggplot(aes(x=(error),y=id,col=Model)) + 
#   stat_pointinterval(position=position_dodge(.2)) + 
#   geom_vline(xintercept = 0, linetype="dashed") +
#   ggh4x::facet_nested_wrap(condit~x, axes = "all",ncol=6,scale="free") +
#   labs(y="Velocity Band", x="Residual Error")
#

```{r  fig.width=11, fig.height=13}
#| label: fig-htw-indv-pred
#| fig-cap: Individual predictions
#| fig-height: 15
#| fig-width: 12

# ids <- c(1,36,66)
# cId <- c(11,37,93,49)
# vId <- c(1,14,193)

cId_tr <- c(137, 181, 11)
vId_tr <- c(14, 193, 47)

cId_tt <- c(11, 93, 35)
vId_tt <- c(1,14,74)

# filter(id %in% (filter(bestTestEXAM,group_rank<=9, Fit_Method=="Test")

post_dat_l |> filter(id %in% c(cId_tr,vId_tr), Fit_Method=="Train",rank<7) |> 
   mutate(x=as.factor(x), Resp=as.factor(Resp)) |>
  ggplot(aes(x = Resp, y = val, fill=x)) + 
  stat_bar + ggh4x::facet_nested_wrap(condit~id, axes = "all",ncol=3)

post_dat_l |> filter(id %in% c(cId_tt,vId_tt), Fit_Method=="Test_Train") |> 
   mutate(x=as.factor(x), Resp=as.factor(Resp)) |>
  ggplot(aes(x = Resp, y = val, fill=x)) + 
  stat_bar + ggh4x::facet_nested_wrap(condit~id, axes = "all",ncol=3)

# post_dat_l |> filter(id %in% (filter(bestTrainEXAM,group_rank<=9,, Fit_Method=="Train") |> pull(id))) |> 
#    mutate(x=as.factor(x), Resp=as.factor(Resp)) |>
#   ggplot(aes(x = Resp, y = val, fill=x)) + 
#   stat_bar + ggh4x::facet_nested_wrap(condit~id, axes = "all",ncol=3)

### Model Comparison
```{r fig.width=11, fig.height=13}
#| eval: false
#| include: false
best_id_x <- post_dat_avg |> group_by(id,condit,Model,Fit_Method,x) |>
  mutate(e2=(y-pred)) |>
  summarise(y=mean(y), pred=mean(pred), mean_error=mean(e2),abs_me=abs(mean_error)) |>
  group_by(id,condit,Fit_Method,x) |> mutate(best=ifelse(mean_error==min(mean_error),1,0)) |>
  group_by(id,condit,Fit_Method,Model) |> mutate(n_best=sum(best))

best_id <- best_id_x |> group_by(id,condit,Fit_Method,Model) |>
  summarise(mean_error=mean(mean_error), n_best=first(n_best),abs_me=mean(abs_me))

full_comparison |> filter(Fit_Method=="Test_Train") |> 
  ungroup() |>
  mutate(id = reorder(id, abs_error_dif)) %>%
  ggplot(aes(y=id,x=abs_error_dif,fill=Best_Model2))+
  geom_col()+
  ggh4x::facet_grid2(~condit,axes="all",scales="free_y", independent = "y")

@@

# full_comparison |> filter(Fit_Method=="Test_Train") |> 
#   ungroup() |>
#   mutate(id = reorder(id, Error_difference)) %>%
#   ggplot(aes(y=id,x=Error_difference,fill=Best_Model))+
#   geom_col()+
#   ggh4x::facet_grid2(~condit,axes="all",scales="free_y", independent = "y")

@@

```{r}
#| label: fig-htw-post-dist
#| fig-cap: Posterior Distributions of $c$ and $lr$
#| fig-height: 10
#| fig-width: 11

c_post <- post_dat_avg %>%
    group_by(id, condit, Model, Fit_Method, rank) %>%
    slice_head(n = 1) |>
    ggplot(aes(y=Fit_Method, x = log(c),col=condit)) + stat_pointinterval(position=position_dodge(.2)) +
    ggh4x::facet_nested_wrap(~Model,scales="free")

lr_post <- post_dat_avg %>%
    group_by(id, condit, Model, Fit_Method, rank) %>%
    slice_head(n = 1) |>
    ggplot(aes(y=Fit_Method, x = lr,col=condit)) + stat_pointinterval(position=position_dodge(.2)) +
    ggh4x::facet_nested_wrap(~Model,scales="free")

c_post + lr_post

@@

>>>>

---
title: HTW Model
author: Thomas Gorman
date: "`r Sys.Date()`"
page-layout: full
code-fold: true
code-tools: true
execute: 
  warning: false
  eval: true
---

```{r}
pacman::p_load(dplyr,purrr,tidyr,ggplot2, data.table, here, patchwork, conflicted, 
               stringr,future,furrr, knitr, reactable, flextable,ggstance, htmltools)
#conflict_prefer_all("dplyr", quiet = TRUE)
options(scipen = 999)
walk(c("Display_Functions","fun_alm","fun_indv_fit","fun_model"), ~ source(here::here(paste0("Functions/", .x, ".R"))))
```

```{r}
#| label: fig-alm-diagram
#| fig.cap: The basic structure of the ALM model. 
 alm_plot()
```

@@

The general ABC procedure is to 
1) define a prior distribution over model parameters.
2) sample candidate parameter values, $\theta^*$, from the prior.
3) Use $\theta^*$ to generate a simulated dataset, $Data_{sim}$.
4) Compute a measure of discrepancy between the simulated and observed datasets, $discrep$($Data_{sim}$, $Data_{obs}$). 
5) Accept $\theta^*$ if the discrepancy is less than the tolerance threshold, $\epsilon$, otherwise reject $\theta^*$.
6) Repeat until desired number of posterior samples are obtained.

Although simple in the abstract, implementations of ABC require researchers to make a number of non-trivial decisions as to i) the discrepancy function between observed and simulated data, ii) whether to compute the discrepancy between trial level data, or a summary statistic of the datasets, iii) the value of the minimum tolerance $\epsilon$. For the present work, we follow the guidelines from published ABC tutorials [@turnerTutorialApproximateBayesian2012; @farrellComputationalModelingCognition2018]. For the test stage, we summarized datasets with mean velocity of each band in the observed dataset as $V_{obs}^{(k)}$ and in the simulated dataset as $V_{sim}^{(k)}$, where $k \in \{1, 2, 3, 4, 5, 6\}$ represents each of the six velocity bands. For computing the discrepancy between datasets in the training stage, we aggregated training trials into three equally sized blocks (separately for each velocity band in the case of the varied group). After obtaining the summary statistics of the simulated and observed datasets, the discrepancy was computed as the mean of the absolute difference between simulated and observed datasets (@eq-discrep-test and @eq-discrep-train).

::: column-page-inset-left
$$
discrep_{Test}(Data_{sim}, Data_{obs}) = \frac{1}{6} \sum_{k=1}^{6} |V_{obs}^{(k)} - V_{sim}^{(k)}|
$${#eq-discrep-test}

$$
\begin{aligned} \\
discrep_{Train,constant}(Data_{sim}, Data_{obs}) = \frac{1}{N_{blocks}} \sum_{j=1}^{N_{blocks}} |V_{obs,constant}^{(j)} - V_{sim,constant}^{(j)}| \\ \\
discrep_{Train,varied}(Data_{sim}, Data_{obs}) = \frac{1}{N_{blocks} \times 3} \sum_{j=1}^{N_{blocks}} \sum_{k=1}^{3} |V_{obs,varied}^{(j,k)} - V_{sim,varied}^{(j,k)}|
\end{aligned}
$${#eq-discrep-train}
:::

@@

```{r}
#| cache: true

ds <- readRDS(here::here("data/e1_md_11-06-23.rds"))  |> as.data.table()
nbins <- 3

@@

input_layer <<- output_layer <<-  c(100,350,600,800,1000,1200)
ids2 <- c(1,66,36)
file_name <- "n_iter_200_ntry_300_5354"

ind_fits <- map(list.files(here(paste0('data/abc_reject/'),file_name),full.names=TRUE), readRDS)
ind_fits_df <- ind_fits |> map(~list(dat=.x[[1]], Model = .x[["Model"]], Fit_Method=.x[["Fit_Method"]]))
ind_fits_df <- ind_fits_df |> map(~rbindlist(.x$dat) |> mutate(Model = .x$Model, Fit_Method = .x$Fit_Method)) |> rbindlist() 
```

```{r}
#| cache: true

@@

post_dat_avg <- post_dat |> group_by(id, condit, Model, Fit_Method, x, c, lr, rank) |> 
  mutate(error2 = y - pred) |>
  summarise(y = mean(y), pred = mean(pred), error = y - pred, error2=mean(error2)) |> as.data.table()

@@

# organize training data predictions
pd_train <- nestSbjModelFit |> mutate(pp=furrr::future_pmap(list(id,Model,Fit_Method,data), ~{
   generate_data(..2, ..4 |> mutate(id=..1), ds, num_samples = 20, return_dat="train_data")
   })) |>
  select(Fit_Method,pp,-data) |>
  unnest(pp) |> as.data.table() |> filter(expMode2=="Train")
plan(sequential)

@@

@@

### Group level fits

```{r}
#| eval: true
#| label: tbl-htw-modelError
#| tbl-cap: "Mean model errors predicting testing data, aggregated over all participants and velocity bands. Note that Fit Method refers to how model parameters were optimized, while error values reflect mean absolute error for the 6 testing bands"
post_tabs <- abc_tables(post_dat,post_dat_l)
post_tabs$agg_pred_full |> flextable::tabulator(rows=c("Fit_Method","Model"), columns=c("condit"), 
                       `ME` = as_paragraph(mean_error)) |> as_flextable() 
```

The posterior distributions of the $c$ and $lr$ parameters are shown @fig-htw-post-dist (i.e. these plots combine all the posterior samples from all of the subjects). There were substantial individual differences in the posteriors of both parameters, with the within-group individual differences generally swamped any between-group or between-model differences. The magnitude of these individual differences remains even if we consider only the single best parameter set for each subject.

We used the posterior distribution of $c$ and $lr$ parameters to generate a posterior predictive distribution of the observed data for each participant, which then allows us to compare the empirical data to the full range of predictions from each model. Model residuals are shown in the upper panels of @fig-htw-resid-pred. The pattern of training stage residual errors are unsurprising across the combinations of models and fitting method . Differences between ALM and EXAM are generally minor (the two models have identical learning mechanisms). The differences in the magnitude of residuals across the three fitting methods are also straightforward, with massive errors for the 'fit to Test Only' model, and the smallest errors for the 'fit to train only' models. It is also noteworthy that the residual errors are generally larger for the first block of training, which is likely due to the initial values of the ALM weights being unconstrained by whatever initial biases participants tend to bring to the task. Future work may explore the ability of the models to capture more fine grained aspects of the learning trajectories. However for the present purposes, our primary interest is in the ability of ALM and EXAM to account for the testing patterns while being constrained, or not constrained, by the training data. All subsequent analyses and discussion will thus focus on the testing stage.

The residuals of model predictions for the testing stage (@fig-htw-resid-pred) show the opposite pattern of fitting method - with models fit only to the test data showing the best performance, followed by models fit to both training and test data, and with models fit only to the training data showing the worst performance (note that y axes are scaled different between plots). Unsurprisingly, the advantage of EXAM is strongest for extrapolation positions (the three smallest bands for both groups - as well as the two highest bands for the Constant group). Although EXAM tends to perform better for both Constant and Varied participants (see also @tbl-htw-modelError), the relative advantage of EXAM is generally larger for the Constant group - a pattern consistent across all three fitting methods. Panel B of @fig-htw-resid-pred directly compares the aggregated observed data to the posterior predictive distributions for the testing stage. Of interest are a) the extent to which the median estimates of the ALM and EXAM posteriors deviate from the observed medians for each velocity band; b) the ability of ALM and EXAM to discriminate between velocity bands; c) the relative performance of models that are constrained by the training data (i.e. the 'fit to train only' and 'fit to both' models) compared to the 'fit to test only' models; and d) the extent to which the variance of the posterior predictive distributions mimics the variance of the observed data.

- ** explain how the constant group ALM predictions for band 100 look deceptively good due to aggregation of a large subset of subjects having ALM predictions of 0 for vb100, and a large subset with ALM predictions close to their position 800 value. This is relected by much greater variance of the ALM esimates in the posterior predictive plot

- ** comment on how much constrained by the training data has a worse impact on the EXAM predictions for varied than for constant - perhaps due to the varied training data being much noisier than the constant training data.

- ** comment on EXAM doing a better job mimicing the within-condition  variance of the observed data
- ** comment on the % of Constant subjects being best accounted for by EXAM being higher.

```{r fig.height=12,fig.width=11}
#| label: fig-htw-resid-pred
#| column: page-inset-right
#| fig-cap: A) Model residuals for each combination of training condition, fit method, and model. Residuals reflect the difference between observed and predicted values. Lower values indicate better model fit. Note that y axes are scaled differently between facets. B) Full posterior predictive distributions vs. observed data from participants.Points represent median values, thicker intervals represent 66% credible intervals and thin intervals represent 95% credible intervals around the median. 
#| fig-height: 19
#| fig-width: 12
##| layout: [[45,-5, 45], [100]]
##| fig-subcap: ["Model Residuals - training data", "Model Residuals - testing data","Full posterior predictive distributions vs. observed data from participants."]
train_resid <- pd_train |> group_by(id,condit,Model,Fit_Method, Block) |> 
  summarise(y = mean(y), pred = mean(pred), error = y - pred) |>
  ggplot(aes(x = Block, y = abs(error), fill=Model)) + 
  stat_bar + 
  ggh4x::facet_nested_wrap(rename_fm(Fit_Method)~condit, scales="free",ncol=2) +
  scale_fill_manual(values=wes_palette("AsteroidCity2"))+
  labs(title="Model Residual Errors - Training Stage", y="RMSE", x= "Training Block") +
  theme(legend.title = element_blank(), legend.position="top")

test_resid <-  post_dat |> 
   group_by(id,condit,x,Model,Fit_Method,rank) |>
   summarize(error=mean(abs(y-pred)),n=n()) |>
   group_by(id,condit,x,Model,Fit_Method) |>
   summarize(error=mean(error)) |>
  mutate(vbLab = factor(paste0(x,"-",x+200))) |>
  ggplot(aes(x = vbLab, y = abs(error), fill=Model)) + 
  stat_bar + 
  scale_fill_manual(values=wes_palette("AsteroidCity2"))+
  ggh4x::facet_nested_wrap(rename_fm(Fit_Method)~condit, axes = "all",ncol=2,scale="free") +
  labs(title="Model Residual Errors - Testing Stage",y="RMSE", x="Velocity Band") +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5))

group_pred <- post_dat_l |> 
  mutate(vbLab = factor(paste0(x,"-",x+200),levels=levels(testAvg$vb))) |>
  ggplot(aes(x=val,y=vbLab,col=Resp)) + 
  stat_pointinterval(position=position_dodge(.5), alpha=.9) + 
  scale_color_manual(values=wes_palette("AsteroidCity2"))+
  ggh4x::facet_nested_wrap(rename_fm(Fit_Method)~condit, axes = "all",ncol=2,scale="free") +
  labs(title="Posterior Predictions - Testing Stage",y="Velocity Band (lower bound)", x="X Velocity") +
theme(legend.title=element_blank(),axis.text.y = element_text(angle = 45, hjust = 0.5, vjust = 0.5))

((train_resid | test_resid) / group_pred) +
  plot_layout(heights=c(1,1.5)) & 
  plot_annotation(tag_levels = list(c('A1','A2','B')),tag_suffix = ') ') & 
  theme(plot.tag.position = c(0, 1))

@@

```{r}
#| label: fig-htw-post-dist
#| fig-cap: Posterior Distributions of $c$ and $lr$ parameters. Points represent median values, thicker intervals represent 66% credible intervals and thin intervals represent 95% credible intervals around the median. Note that the y axes of the plots for the c parameter are scaled logarithmically.
#| fig-height: 10
#| fig-width: 11

c_post <- post_dat_avg %>%
    group_by(id, condit, Model, Fit_Method, rank) %>%
    slice_head(n = 1) |>
    ggplot(aes(y=log(c), x = Fit_Method,col=condit)) + stat_pointinterval(position=position_dodge(.2)) +
    ggh4x::facet_nested_wrap(~Model) + labs(title="c parameter") +
  theme(legend.title = element_blank(), legend.position="right",plot.title=element_text(hjust=.4))

lr_post <- post_dat_avg %>%
    group_by(id, condit, Model, Fit_Method, rank) %>%
    slice_head(n = 1) |>
    ggplot(aes(y=lr, x = Fit_Method,col=condit)) + stat_pointinterval(position=position_dodge(.2)) +
    ggh4x::facet_nested_wrap(~Model) + labs(title="learning rate parameter") +
  theme(legend.title = element_blank(), legend.position = "none",plot.title=element_text(hjust=.4))
c_post + lr_post
```

To more accurately assess the relative abilities of ALM and EXAM to capture important empirical patterns - we will now examine the predictions of both models for the subset of individual participants shown in @fig-htw-indv-pred. Panel A presents three varied and constant participants who demonstrated a reasonable degree of discrimination between the 6 velocity bands during testing.

- ** comment on the different ways ALM can completely fail to mimic discrimination patterns (sbj. 35; sbj. 137),and on how it can sometimes partially succeed (sbj. 11; 14,74)

- ** comment on how EXAM can somtimes mimic non-monotonic spacing between bands due to associative stregth from training (i.e. subject 47)

- ** compare c values to slope parameters from the statistical models earlier in paper

```{r  fig.width=11, fig.height=10}
#| label: fig-htw-indv-pred
#| fig-cap: Model predictions alongside observed data for a subset of individual participants. A) 3 constant and 3 varied participants fit to both the test and training data. B) 3 constant and 3 varied subjects fit to only the trainign data. 
#| fig-height: 16
#| fig-width: 12

cId_tr <- c(137, 181, 11)
vId_tr <- c(14, 193, 47)
cId_tt <- c(11, 93, 35)
vId_tt <- c(1,14,74)
# filter(id %in% (filter(bestTestEXAM,group_rank<=9, Fit_Method=="Test")

testIndv <- post_dat_l |> filter(id %in% c(cId_tt,vId_tt), Fit_Method=="Test_Train") |> 
   mutate(x=as.factor(x), Resp=as.factor(Resp)) |>
  group_by(id,condit,Fit_Method,Model,Resp) |>
   mutate(flab=paste0("Subject: ",id)) |>
  ggplot(aes(x = Resp, y = val, fill=x)) + 
  stat_bar_sd + ggh4x::facet_nested_wrap(condit~flab, axes = "all",ncol=3) +
  labs(title="Individual Participant fits from Test & Train Fitting Method",
       y="X Velocity",fill="Target Velocity") +
   guides(fill = guide_legend(nrow = 1)) + 
  theme(legend.position = "bottom",axis.title.x = element_blank())

trainIndv <- post_dat_l |> filter(id %in% c(cId_tr,vId_tr), Fit_Method=="Train") |> 
   mutate(x=as.factor(x), Resp=as.factor(Resp), flab=paste0("Subject: ",id)) |>
  group_by(id,condit,Fit_Method,Model,Resp) |>
  ggplot(aes(x = Resp, y = val, fill=x)) + 
  stat_bar + 
  ggh4x::facet_nested_wrap(condit~flab, axes = "all",ncol=3) +
  labs(title="Individual Participant fits from Train Only Fitting Method", y="X Velocity",
       fill="Target Velocity") +
     guides(fill = guide_legend(nrow = 1)) + 
  theme(legend.position = "bottom",axis.title.x = element_blank())

(testIndv  / trainIndv) +
  plot_annotation(tag_levels = list(c('A','B')),tag_suffix = ') ') & 
  theme(plot.tag.position = c(0, 1))

@@

# 
# 
# full_comparison |> filter(Fit_Method=="Test_Train") |>
#   ungroup() |>
#   mutate(id = reorder(id, Error_difference)) %>%
#   ggplot(aes(y=id,x=Error_difference,fill=Best_Model))+
#   geom_col()+
#   ggh4x::facet_grid2(~condit,axes="all",scales="free_y", independent = "y")
#

@@

@@

### To add to appendix

```{r fig.width=11, fig.height=10}
#| label: fig-htw-best-model
#| fig-cap: Difference in model errors for each participant, with models fit to both train and test data. Positive values favor EXAM, while negative values favor ALM. 
#| fig-height: 10
#| fig-width: 11
best_id_x <- post_dat_avg |> group_by(id,condit,Model,Fit_Method,x) |>
  mutate(e2=(y-pred)) |>
  summarise(y=mean(y), pred=mean(pred), mean_error=mean(e2),abs_me=abs(mean_error)) |>
  group_by(id,condit,Fit_Method,x) |> mutate(best=ifelse(mean_error==min(mean_error),1,0)) |>
  group_by(id,condit,Fit_Method,Model) |> mutate(n_best=sum(best))

best_id <- best_id_x |> group_by(id,condit,Fit_Method,Model) |>
  summarise(mean_error=mean(mean_error), n_best=dplyr::first(n_best),abs_me=mean(abs_me))

full_comparison |> filter(Fit_Method=="Test_Train") |> 
  ungroup() |>
  mutate(id = reorder(id, abs_error_dif)) %>%
  ggplot(aes(y=id,x=abs_error_dif,fill=Best_Model2))+
  geom_col()+
  ggh4x::facet_grid2(~condit,axes="all",scales="free_y", independent = "y")+
  labs(fill="Best Model",x="Mean Model Error Difference (ALM - EXAM)",y="Participant")
